---
documentclass: jss
author:
  - name: FirstName LastName
    affiliation: University/Company
    address: >
      First line
      Second line
    email: \email{name@company.com}
    url: https://github.com/mages/ChainLadder
  - name: Second Author
    affiliation: Affiliation
title:
  formatted: "Claims reserving with R: \\pkg{ChainLadder}"
  # If you use tex in the formatted title, also supply version without
  plain:     "Claims reserving with R: ChainLadder"
  # For running headers, if needed
  short:     "Claims reserving with R: \\pkg{ChainLadder}"
abstract: >
  The ChainLadder package provides various statistical methods
  which are typically used for the estimation of outstanding claims reserves
  in general insurance, including those to estimate the claims development
  results as required under Solvency II.
keywords:
  # at least one keyword must be supplied
  formatted: [reserving, non-life insurance, "\\proglang{R}"]
  plain:     [reserving, non-life insurance, R]
preamble: >
  \usepackage{amsmath}
output: rticles::jss_article
---

\pagebreak

# Introduction

\pagebreak

# Package Overview
## Existing tools
## Our Package

\pagebreak

# Support functions for data manipulation
## Working with triangles

Historical insurance data is often presented in form of a triangle structure, showing
the development of claims over time for each exposure (origin) period. An origin
period could be the year the policy was written or earned, or the loss occurrence
period. Of course the origin period doesn’t have to be yearly, e.g. quarterly or
6 monthly origin periods are also often used. The development period of an origin
period is also called age or lag. Data on the diagonals present payments in the
same calendar period. Note, data of individual policies is usually aggregated to
homogeneous lines of business, division levels or perils.

### Creating triangles interactively

We can create a triangle object interactively using R vectors and the `triangle` function provided by the package:

```{r message=FALSE}
require(ChainLadder)
```
```{r}
triangle(c(100, 150, 175, 180, 200),
         c(110, 168, 192, 205),
         c(115, 169, 202),
         c(125, 185),
         150)
```

We can optionally specify the *labels* for the origin periods:
```{r}
triangle("2012" = c(100, 150, 175, 180, 200),
         "2013" = c(110, 168, 192, 205),
         "2014" = c(115, 169, 202),
         "2015" = c(125, 185),
         "2016" = 150)
```

and we can also specify the labels for development periods:
```{r}
triangle("2012" = c("12 months" = 100,
                    "24 months" = 150,
                    "36 months" = 175,
                    "48 months" = 180,
                    "60 months" = 200),
         "2013" = c(110, 168, 192, 205),
         "2014" = c(115, 169, 202),
         "2015" = c(125, 185),
         "2016" = 150)
```

If a single vector is passed to the triangle constructor, the function will try to infer the structure of the desired triangle:
```{r}
triangle(c(100, 150, 175, 110, 168, 115))
```

### Converting data into triangles
### Visualizing triangles
### Integration with Excel

\pagebreak

# ChainLadder: Models and Methods
## Classic chain-ladder methods
### Mack chain-ladder
### BootChainLadder
## Growth curves
### Clark LDF
## Modelling paid and incurred data
### MunichChainLadder
### Paid-incurred chain model
## Multivariate chain-ladder
## GLMs
### Tweedie model

\pagebreak

# Solvency II: Claims Development Result
## CDR function

\pagebreak

# Support functions for data manipulation and model validation
## Interpreting graphical outputs
## Model validation with tweedieReserve

\pagebreak

# Conclusion

\pagebreak
